---
title: "Assignment"
output: html_document
date: "2025-11-12"
---
## 1. 
Write a function to calculate the standard deviation from the population variance of
the dataset `pressure`, for the field `pressure`. Display the output along with the values
of this field.

```{r calculate-sd-from-var, echo=TRUE}
calculate_sd_from_var <- function(data, field) {
  values <- data[[field]]
  
  # Calculate population variance
  pop_variance <- sum((values - mean(values))^2) / length(values)
  
  # Calculate population standard deviation
  pop_sd <- sqrt(pop_variance)
  
  # Display results
  cat("Field:", field, "\n")
  cat("Values:\n")
  print(values)
  cat("\nPopulation Variance:", pop_variance, "\n")
  cat("Standard Deviation (from population variance):", pop_sd, "\n")
  
  # Return SD value
  return(pop_sd)
}

# Example usage
data(pressure)
calculate_sd_from_var(pressure, "pressure")
```

## 2.
Take the dataset ‘Titanic’. List out the categorical fields and create cross-tabulations
of Sex vs Survived and Age vs Survived, along with the chi-square proportions and
row and column totals listed out.

# Load and prepare dataset
```{r titanic_setup}
data(Titanic)
titanic_df <- as.data.frame(Titanic)
```
# Cross-tabulation: Sex vs Survived
tab_sex_surv <- xtabs(Freq ~ Sex + Survived, data = titanic_df)
tab_sex_surv

# Add row and column totals
addmargins(tab_sex_surv)

# Chi-square test
chi_sex_surv <- chisq.test(tab_sex_surv)
chi_sex_surv

# Chi-square proportions (expected proportions)
chi_sex_surv$expected

# Cross-tabulation: Age vs Survived
tab_age_surv <- xtabs(Freq ~ Age + Survived, data = titanic_df)
tab_age_surv

# Add row and column totals
addmargins(tab_age_surv)

# Chi-square test
chi_age_surv <- chisq.test(tab_age_surv)
chi_age_surv

# Expected proportions
chi_age_surv$expected

## 3.
Differentiate between sample and population variance, along with examples and a
demonstration with an actual dataset.


